\Title

### 2022-09-13 Grimaldi-etal-BiolCybernetics.tex
task

%
%\ninept
%
%\abstract{Abstract}
%
\keywords{\Keywords}
%
\maketitle
%%%%%%%%% BODY TEXT
\section{Introduction}
\label{sec:intro}
%
In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\citep{abeles_role_1982}. While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code. Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}. All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in biological spike trains.

In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}. Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays. This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
However, most current neuroscience-inspired computer vision algorithms (for instance convolution neural networks) do not make use of this dynamic aspect. A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance. The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations. Yet, flexible and robust event-driven algorithms for classical computer vision tasks, such as motion detection, are still not able to compete with the state-of-the-art of dense computer vision solutions.

In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motionclouds}). We develop a SNN-like method able to learn hetero-synaptic delays to perform motion detection on a synthetic event-based dataset. Because neuromorphic devices are, by design, good candidates to integrate computations with time, we highlight the fact that this event-driven algorithm is transferable to such hardware.
%
\section{Methods}
\label{sec:methods}
%
In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}. This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events. By transforming each event in the stream as a vectorial input, our MLR classifier is able to make a decision for every single event, that is in an online fashion. We have demonstrated on several datasets that it provides with efficient computations resulting in ultra-fast categorization. Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.

Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights. Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays. In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay as a computational process. As a consequence, this extends our previous model by encoding spiking patterns as 3D images. The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier. In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these hetero-synaptic delays. %
%
We expect our model to be able to efficiently detect temporal patterns and here we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion.
%
\subsection{Task definition: fast motion detection}
%
Taking analogy with biological conditions, a common use case for motion detection is to estimate the motion generated on a sensor by saccade-like movements. Saccades are rapid movements of the eyeball which allow to reposition the axis of view in a short period. In humans, these are particularly rapid 
To test our model, we will quantify its ability to categorize different motions. In that order, we will generate a natural

@@

\subsection{Hetero-synaptic delays model}
%
\begin{figure*}[ht!]
    {\centering
    \vspace{-3cm}
    \includegraphics[width=\linewidth]{figures/3D_kernels_oneline.png}
    }
    %\vspace{-.5cm}
    \caption{
    	Representation of the weights for the $8$ learned kernels of the model corresponding to the OFF polarities and selective to the different motion directions
	(because of the symmetry observed between the ON and OFF event streams, kernels are similar for the ON polarities). These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown. One sees positive (excitatory) coefficients for the specific direction of motion and negative (inhibitory) coefficients for all other directions.
	}
    \label{fig:kernels}
\end{figure*} 
%
The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$. 
This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).

@@

\subsection{
Temporal
Logistic Regression}

>>>>

%
%\ninept
%
%\abstract{Abstract}
%
\keywords{\Keywords}
%
\maketitle
%%%%%%%%% BODY TEXT
\section{Introduction}
\label{sec:intro}
%
\subsection{Dynamics of sensory input and spiking neurons}
In 1982, Abeles asked if the role of cortical neurons is whether to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns~\citep{abeles_role_1982}. While the first possibility favors the rate coding theory, the second highlights the function of temporal precision in the neural code. Since, numerous studies demonstrated the emergence of synchronicity in the activity within a neural population~\citep{riehle_spike_1997, davis_spontaneous_2021}, efficient encoding thanks to the use of spike latencies~\citep{perrinet_coding_2004, gollisch_rapid_2008} or precise timing in the auditory system~\citep{deweese_binary_2002, carr_circuit_1990}. All these findings, and more~\citep{bohte_evidence_2004}, highlight the importance of the temporal aspect of the neural code and suggest the existence of repeated spatio-temporal patterns in biological spike trains.

\subsection{Hetero-synaptic delays}
In neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{gutig_tempotron_2006, guise_bayesian_2014, zhang_supervised_2020}. Notably,~\citet{izhikevich_polychronization_2006} introduced the notion of the polychronous group as a repetitive motif of spikes defined by a subset of neurons with different, yet precise, relative spiking delays. This representation has a much greater information capacity in comparison to a firing-rate based neural coding approach through the variety of configurations and the possible coexistence of multiple superposed motifs.
However, most current neuroscience-inspired computer vision algorithms (for instance convolution neural networks) do not make use of this dynamic aspect. A novel emerging representation is that provided by event-based cameras, in which each pixel independently processes its input and emits an event for positive or negative increments of the log-luminance. The shift from the classical, dense representations to this sparse encoding of visual information offers a better analogy to neurobiology, but also offers more energy-efficient computations. Yet, flexible and robust event-driven algorithms for classical computer vision tasks, such as motion detection, are still not able to compete with the state-of-the-art of dense computer vision solutions.

\subsection{Outline}

In a recent study, we have introduced a spiking neural network as a classification model applied to event streams based on Multinomial Logistic Regression (MLR)~\citep{grimaldi_robust_2022} which extended the HOTS model~\citet{lagorce_hots_2017}. This HOTS model represents temporal relationships between events by building ``time surfaces'', 2D images computed using the time difference to the last recorded events. By transforming each event in the stream as a vectorial input, our MLR classifier was able to make a decision for every single event, that is in an online fashion. We have demonstrated on several datasets that it provides with efficient computations resulting in ultra-fast categorization. Additionally, we made a formal bridge between this event-based MLR and a SNN, demonstrating the bio-plausibility of this method and its possible integration to neuromorphic hardware.

Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights. Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, hetero-synaptic delays. In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay in the computational process. As a consequence, this extends our previous model by encoding spiking motifs as 3D images. The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier. In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these hetero-synaptic delays. %

In this work, we study the emergence of such spatio-temporal spiking motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motionclouds}). We develop a SNN-like method able to learn hetero-synaptic delays to perform motion detection on a synthetic event-based dataset. Because neuromorphic devices are, by design, good candidates to integrate computations with time, we highlight the fact that this event-driven algorithm is transferable to such hardware.
%
\section{Methods}
\label{sec:methods}
%
%
We expect our model to be able to efficiently detect temporal patterns and here we will challenge it by defining a realistic task of motion detection in a stream generated by a natural scene with known ground truth motion.
%
\subsection{Task definition: fast motion detection}
%
Taking analogy with biological conditions, a common use case for motion detection is to estimate the motion generated on a sensor by saccade-like movements. Saccades are rapid movements of the eyeball which allow to reposition the axis of view in a short period. In humans, these are particularly rapid 
To test our model, we will quantify its ability to categorize different motions. In that order, we will generate a natural

@@

\subsection{Hetero-synaptic delays model as a temporal Logistic Regression}
%
\begin{figure*}[ht!]
    {\centering
    \vspace{-3cm}
    \includegraphics[width=\linewidth]{figures/3D_kernels_oneline.png}
    }
    %\vspace{-.5cm}
    \caption{
    	Representation of the weights for the $8$ learned kernels of the model corresponding to the OFF polarities and selective to the different motion directions
	(because of the symmetry observed between the ON and OFF event streams, kernels are similar for the ON polarities). These weights are associated to a specific delay on the \textit{delays} axis and to a presynaptic address defined on the two other axes.
	For the sake of clarity, the values in range $[-0.05, 0.05]$ are not shown. One sees positive (excitatory) coefficients for the specific direction of motion and negative (inhibitory) coefficients for all other directions.
	}
    \label{fig:kernels}
\end{figure*} 
%
The sensory signal representing the output of an event-based camera forms a discrete stream of events, which can be formalized as an ordered set of addresses and timestamps: $\event = \{(\presynaddr_\arank, \timev_\arank)\}_{\arank \in [1,\numevent]}$ where $\numevent \in \mathbb{N} $ is the total number of events in the data stream and the rank $\arank$ is the index of each event $\event_\arank$. 
This event has a time of occurrence $\timev_\arank$  and an associated address, which is typically in the form $\presynaddr_\arank=(x_\arank, y_\arank, \polev_\arank)$.
This defines a presynaptic address space $\presynaddrspace = [1, \Nx]\times[1, \Ny] \times [1, \Npol] \subset \mathbb{N}^3$ where $(\Nx, \Ny)$ is the size of the sensor in pixels and $\Npol$ is the number of polarities  ($\Npol=2$ for ON and OFF polarities).

@@

\subsection{A Spiking Neural Network detecting motion as a CNN}
